---
title: "Cryptocurrency Analysis"
author: "Damien Fleminks"
date: "21/11/2020"
output: 
  html_document: 
    fig_width: 9
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### install and setup packages


```{r loadlib, echo=T, results='hide', message=F, warning=F}
#  rvest for web scraping
#  -> install.packages('rvest')
library(rvest)

#  tidyverse & gridExtra for data manipulation & visualization
#  -> install.packages("tidyverse")
#  -> install.packages("gridExtra")
library(tidyverse)
library(gridExtra)

# httr & jsonlite for rest api
#  -> install.packages("httr")
#  -> install.packages("jsonlite")
require("httr")
require("jsonlite")

#  rlist to extend list functions
#  -> install.packages("rlist")
library(rlist)
```

### Web scraping and data cleanup


Use coinranking.com to get top 50 cryptocurrency coins and retrieve the ticker symbol, coin name & current market cap

```{r}
crypto.url <- read_html("https://coinranking.com/")

crypto.tickers <- crypto.url %>% 
  html_nodes(".profile__subtitle") %>% 
  html_text()

crypto.name <- crypto.url %>% 
  html_nodes(".profile__link") %>% 
  html_text()

crypto.market.cap<- crypto.url %>% 
  html_nodes(".valuta") %>% 
  html_text()
```


Clean up the strings by removing unused characters and split into separate vectors 

```{r}
crypto.tickers <- str_trim(str_remove_all(crypto.tickers, "\n"))
crypto.name <- str_trim(str_remove_all(crypto.name, "\n"))
crypto.unit.price <- crypto.market.cap[c(TRUE,FALSE)]
crypto.market.cap <- crypto.market.cap[c(FALSE,TRUE)]
```


Convert to correct data types and track which market cap price uses price unit in billion so we can multiply all output into units of million

```{r}
crypto.isbil <-  str_detect(crypto.market.cap,"billion")
crypto.unit.price <-  as.numeric(gsub("[^0-9.]", "", crypto.unit.price))

# find billion values and change them into million units
crypto.market.cap <- as.numeric(gsub("[^0-9.]", "", crypto.market.cap))
for(i in 1:length(crypto.isbil)){
     if(crypto.isbil[i]){
         crypto.market.cap[i] <- crypto.market.cap[i] * 1000
     }
}

# Create data frame of scraped and formatted data
crypto.currencies <- data.frame(Cryptocurrency = crypto.name, Ticker = crypto.tickers, Price=crypto.unit.price, market.cap = crypto.market.cap)
```


We now have a cleaned up Data frame with currency data that was taken from coinranking.com. We will use this data to retrieve historical price data from an api.

```{r crypto.currencies}
print(
  head(crypto.currencies)
)
```





### Get historical price data from REST API


With the data collected from the web we will retrieve historical pricing data from api.tiingo.com API, we will start firt by setting up the URL to execute a GET request.


```{r}

#  Setting up base URL for cryptocurrency price history API
api.base <- "https://api.tiingo.com/tiingo/crypto/prices"

# Set currency used for pricing cryptocurrencies
api.currency <- "usd"

#  Retrieve history for top 5 cryptocurrencies, using the tickers from our cryptocurrency data frame
api.tickers <- paste(crypto.currencies$Ticker[1:5],"usd",sep="")
api.tickers <- paste(api.tickers, collapse = ',')
# Start date to retrieve prices in YYYY-MM-DD format, set 1 year from today
api.startdate <- as.POSIXlt(Sys.Date())
api.startdate$year <- api.startdate$year-1
api.startdate <- as.Date(api.startdate)

# set the frequency in which you want data resampled 
api.resampleFreq = "1day"

# API token for authentication  (A free token can be requested at: api.tiingo.com, but you can use mine for now)
api.token = "20b7879009ee4863aef6a4dfd06fb50fb9ea447e"

# Combine variables into GET URL 
api.call <- paste(api.base, "?tickers=",api.tickers,"&startDate=",api.startdate,"&resampleFreq=",api.resampleFreq,"&token=",api.token,sep='')
```


Make GET request with the request URL and store response in variable, we will convert the result into JSON and clean up the data and format it into a usuable format for our analysis.

```{r, echo=T, results='hide', message=F, warning=F}
api.prices <- GET(api.call)

#  Format response to JSON
api.prices_text <- content(api.prices, "text")
api.prices_json <- fromJSON(api.prices_text, flatten = TRUE)
api.prices_df <- as.data.frame(api.prices_json)

#  Create a single data frame with formatted data for easier data visualization and  correct data types

returnNewDf <- function(index){
  
  #  Get coin and store dataframe
  temp.coin <- api.prices_df$baseCurrency[index]
  temp.df <- api.prices_df[index,][[4]][[1]]
  
  # Get length and store coin name as new column
  temp.rownumber <- length(temp.df[,1])
  temp.df$coin <- rep(c(temp.coin),each=length(temp.rownumber))
  temp.df$date <- as.Date(temp.df$date)
  
  #  To calculate the price value of max in terms of percentage add new column with percentages of close 
  temp.df$price.percentage <- temp.df$close/max(temp.df$close)
  
  return(temp.df)
}

#  Add first data frame, then loop to add the rest
crypto.all <- returnNewDf(1)
for(i in 1:length(api.prices_df[,1])) {
  crypto.all <- rbind(
    crypto.all,returnNewDf(i)
    
  )
}

```


Now we can access all coins from the same data frame with the different unique coins, looking at the head of 2 examples


```{r crypto.all}
unique(crypto.all[c('coin')])

print(
  head(filter(crypto.all, coin=='btc'))
)
print(
  head(filter(crypto.all, coin=='eth'))
)
```


### Visualizing the data

We can analyze the price of the coins over time to see how they have performed over the past year.

```{r}

# Plot a line graph for all coin prices
cPrice1 <- ggplot(filter(crypto.all, coin=='btc'), aes(x=date,y=close)) + geom_line() + ggtitle("Bitcoin")
cPrice2 <- ggplot(filter(crypto.all, coin=='eth'), aes(x=date,y=close)) + geom_line() + ggtitle("Ethereum")
cPrice3 <- ggplot(filter(crypto.all, coin=='link'), aes(x=date,y=close)) + geom_line() + ggtitle("Chainlink")
cPrice4 <- ggplot(filter(crypto.all, coin=='xrp'), aes(x=date,y=close)) + geom_line() + ggtitle("XRP")
grid.arrange(cPrice1 , cPrice2, cPrice3, cPrice4, nrow = 2)
```

As we have added an extra column to compare the percentage of the price to the max price we can compare how all coins compared relatively from each other in terms of price fluctation.

```{r}

# Compare the different coins on the percentage growth
ggplot(crypto.all, aes(x=date,y=price.percentage)) + geom_line(aes(color=coin)) + ggtitle("Price Fluctation")      
```


From the tradesDone and volumeNotional column we can analyze the number of transitions and total volume in USD traded, we will compare the results and look at the daily trend over the year and the total numbers.


```{r}
# Compare  frequency & volume in USD of trades over the last year

cTradeComp <- ggplot(crypto.all, aes(x=date,y=tradesDone)) + geom_line(aes(color=coin))  + ggtitle("Transactions per day")  
cVolComp <- ggplot(crypto.all, aes(x=date,y=volumeNotional)) + geom_line(aes(color=coin))  + ggtitle("Volume in USD per day")  
grid.arrange(cTradeComp , cVolComp, nrow = 2)

# Compare total frequency & volume of trade 

cTradeTotal <-ggplot(crypto.all) + geom_bar(aes(y=tradesDone, x=coin, fill=factor(coin)), stat="identity")  + ggtitle("Total transactions")  
cVolTotal <- ggplot(crypto.all) + geom_bar(aes(y=volumeNotional, x=coin, fill=factor(coin)), stat="identity") + ggtitle("Total volume traded in USD")  
grid.arrange(cTradeTotal , cVolTotal, nrow = 1)

```


